#My First knitr Document
=======================

This is some text 

Here is a code chunk
```{r echo =FALSE}
set.seed(1)
x<-rnorm(100)
mean(x)
```

##Inline Text Computations
```{r computrtime, echo=FALSE}
time <- format(Sys.time(),"%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`. My favorite random number is `r rand`

##Incorporating graphics
Let's first simulate some data
```{r simulatedata, echo = TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd=0.5)
```

Here is a scatterplot of the data
```{r scatterplot, fig,height=4}
par(mar=c(5,4,1,1), las=1)
plot(x,y,main="My Simulated Data")
```

##Making Tables with xtable

```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data=airquality)
```

Here is a table of regression coefficients
```{r showtable, results ="asis"}
library(xtable)
xt<-xtable(summary(fit))
print(xt,type="html")
```

##Setting Global Options
```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, results="hide")
```

First simulate data
```{r simulatedata global options, echo=TRUE}
x<-rnorm(100); y<-x +rnorm(100, sd=0.5)
```

Here is a scatterplot of the data
```{r scatterplot,fig.height=4}
par(mar=c(5,4,1,1), las=1)
plot(x,y,main ="My simulated Data")
```