---
title: "Getting and Cleaning Data Project"
author: "Sherifat Akintunde-Shitu"
date: "19 June 2018"
output: 
  html_document:
    keep_md: yes
    toc: yes
---
#Introduction
This file describes the variables, the data, and any transformations or work that you performed to clean up the data for the Getting and Cleaning Data Project. This file is split into three sections as seen in the table of content above. 

The purpose of this project is to demonstrate the ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis.
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(codebook)
```
#Analysis
## *Getting data*
In this section, I 
- download the UCI HAR Dataset from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip)- unzip the zip file and store the data from the files in tbl_dfs with corresponding names. the resulting tbl_dfs are as follow

        - *features*: List of all features.
        - *activity_labels*: Links the class labels with their activity name.
        - *X_train*: Training set.
        - *X_train*: Training labels.
        - *X_test*: Test set.
        - *Y_test*: Test labels.
        - *subject_train*: Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 
        - *subject_train*: Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 

```{r getdata, echo=FALSE}
#download project data set
if(!file.exists("./data")){dir.create("./data");download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", destfile="./data/UCI HAR Dataset.zip" )}

#unzipdata
setwd("./data")
unzip("UCI HAR Dataset.zip")
setwd("../")

#read datasets 
activity_labels<-tbl_df(read.table("./data/UCI HAR Dataset/activity_labels.txt"))
features<-tbl_df(read.table("./data/UCI HAR Dataset/features.txt"))
subject_train<-tbl_df(read.table("./data/UCI HAR Dataset/train/subject_train.txt"))
X_train<-tbl_df(read.table("./data/UCI HAR Dataset/train/X_train.txt"))
Y_train<-tbl_df(read.table("./data/UCI HAR Dataset/train/Y_train.txt"))
subject_test<-tbl_df(read.table("./data/UCI HAR Dataset/test/subject_test.txt"))
X_test<-tbl_df(read.table("./data/UCI HAR Dataset/test/X_test.txt"))
Y_test<-tbl_df(read.table("./data/UCI HAR Dataset/test/Y_test.txt"))
```

##*Data Processing*

###Step 1. Extract Mean and Standard Deviation Measures
To extract the mean and standard deviation measures, I used the grep() function to get IDs of features that have mean() and std() in their text and select subsets from X_test and X_train sets into ms_test and ms_train dataframes.

        - *ms_test* :subset of x_test with mean() and std() measures
        - *ms_train* :subset of x_train with mean() and std() measures
```{r extractms}
# extract mean n std measures for tran and test sets
        ms_test<-select(X_test,grep("mean\\W|std\\W",as.character(features$V2)))
        ms_train<-select(X_train,grep("mean\\W|std\\W",as.character(features$V2)))
```
        
###Step 2. Bind Column labels
To create a more complete data set I bind the test and train subsets with their corresponting subject, activity and lables to form to form the test and train dfs

        - *test* : column bind of subject_test, Y_test and ms_test
        - *train* : column bind of subject_train, Y_train and ms_train
```{r colbind data}
#bind test subject, set and labels
        test<-cbind(subject_test,Y_test,ms_test)
#bind train subject, set and labels
        train<-cbind(subject_train,Y_train,ms_train)
```

###Step 3. Merge the train and test data sets

        - *data* : row bind of test and train data sets along with the activiry, subject and labels

```{r mergedata}
#merge test and train data sets
        data<-rbind(test,train)

```
###Step 4. To set the column names, I 
- extract the mean and standard deviation measures from the feautures df using the grep() function to get valuess of feautures that have mean() and std() in their text.
- use tolower() to make all the text lower case
- concantenate the resulting vector with vector c("subject","activity")
- save the concatenated vector as variables
- removed the '()' in the text using the gsub() function
- and then set the column names od the tbl_df : data to variables
        
        -*variables*: column names for data set
```{r set colname}
#set colnames for merged data set
        variables<-c(c("subject","activity"),
             tolower(grep("mean\\W|std\\W",
                          as.character(features$V2),value=TRUE)))
        variables<-gsub("\\()","", variables)
        colnames(data)<- variables
```

###Step 5. Set Activity Names
To set activity names
- create a list of activity from the activity labels 
- using recode_factor() replace activity IDs with corresponding activity

        - *activity_list* :list of activity and coresponding IDs
        
```{r set activitynames}
#give activities in the data set a descriptive name
        activity_list= as.list(as.character(activity_labels$V2),as.numeric(activity_labels$V1))
        data$activity<-recode_factor(data$activity, !!!activity_list)
```

###Step 6 Create Tidy Data Set with average Values
To create independent tidy data set with the average of each variable for each activity and each subject, 
- group data by activity, subject
- summarise the data using summarise_all() calling the mean function

        - *mean_data* - final tidy data set with the average of each variable for each activity and each subject
```{r average dataset}
        mean_data <-data  %>%
          group_by(activity,subject) %>%
          summarise_all(mean)
```

##*Outputing Result*
Output the tidy data set to a space separated file :'tidydata.txt'
```{r}
#write tidy data set to 
write.table(mean_data, file = "tidydata.txt", sep = " ", row.names = FALSE)
```